Supplement for: Carbon cycling in mature and regrowth forests globally: a macroecological synthesis based on the global Forest Carbon (ForC) database

### 2021-01-23 manuscript/ERL_global_C_review.Rmd
combined stacked area results plots into one, per suggestion of @hmullerlandau
#64

#figures
schematic =1
schematic_fluxes_biome="1a"
  schematic_fluxes_age="1b"
  schematic_stocks_biome="1c"
  schematic_stocks_age="1d"
  schematic_fs_age = "1b,d"
figure_map=2
diagram_TrB_mature= 3
diagram_TeB_mature=4
diagram_TeN_mature=5
diagram_BoN_mature=6
all_diagrams_mature="3-6"
flux_age_trends=7
stock_age_trends=8
TrB_age_trends=9
TeB_age_trends=10
TeN_age_trends=11
BoN_age_trends=12

@@

TrB_age_trends_caption = "Figure 9 | Age trends in C cycling in tropical broadleaf forests. Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle. Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows: $R_{auto-ag}=R_{auto}-R{root}$, where $R_{auto}=NPP(1/CUE-1)$, where $CUE$=0.46 (Collati et al. 2020); $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$. Note that there remain substantial uncertainties as to the functional form of age trends and discrepancies in closure among related variables."
age_trends_caption_generic= "Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle. Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows: $R_{auto-ag}=R_{eco}-R{soil}$; $ANPP_{woody}=min(ANPP_{stem},ANPP_{woody})$; $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$; $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$; $DW_{standing}=max(0, DW_{tot}-DW_{down})$. Note that there remain substantial uncertainties as to the functional form of age trends and discrepencies in closure among related variables."
TeB_age_trends_caption = paste("Figure 10 | Age trends in C cycling in temperate broadleaf forests.", age_trends_caption_generic) 
TeN_age_trends_caption = paste("Figure 11 | Age trends in C cycling in temperate conifer forests.", age_trends_caption_generic) 
BoN_age_trends_caption = paste("Figure 12 | Age trends in C cycling in boreal conifer forests.", age_trends_caption_generic)

@@

C fluxes commonly increased significantly with stand age (Tables 1, S2, Figs. `r flux_age_trends`, `r TrB_age_trends`-`r BoN_age_trends`, `r all_age_trends`).
*ForC* contained 16 C flux variables with sufficient data for analyses of age trends in young forests (see Methods).
Of these, ten increased significantly with $log10[age]$: $NEP$, $GPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{foliage}$, $ANPP_{litterfal}$, $BNPP$, $BNPP_{fine}$, $R_{eco}$, and $R_{root}$.
The remaining six--$NPP$, $ANPP_{stem}$, $ANPP_{branch}$, $BNPP_{coarse}$, $R_{soil}$, and $R_{het-soil}$-–displayed no significant relationship to stand age.

```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =TrB_age_trends_caption, out.width='100%', fig.pos='H'}

  knitr::include_graphics("tables_figures/stacked_age_trends/TrB_age_trends.png")
```

```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =TeB_age_trends_caption, out.width='100%', fig.pos='H'}

  knitr::include_graphics("tables_figures/stacked_age_trends/TeB_age_trends.png")
```

```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =TeN_age_trends_caption, out.width='100%', fig.pos='H'}

  knitr::include_graphics("tables_figures/stacked_age_trends/TeN_age_trends.png")
```

```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =BoN_age_trends_caption, out.width='100%', fig.pos='H'}

  knitr::include_graphics("tables_figures/stacked_age_trends/BoN_age_trends.png")
```

@@

"Closure" and internal consistency of the C flux budget was less successful for young than mature forests (Figs. `r TrB_age_trends`-`r BoN_age_trends`).
Summed regression equations for $R_{soil-het}$ and $R_{root}$ were generally very close to $R_{soil}$.
We calculated $R_{auto-ag}$ as the difference between $R_{eco}$ and $R_{soil}$ (except for tropical forests, which had insufficient $R_{eco}$ data), effectively guaranteeing near-closure of the CO~2~ efflux (respiration) portion of the budget (negative values in Figs. `r TrB_age_trends`-`r BoN_age_trends`).
In contrast, the  CO~2~ influx portion of the budget generally did not "close": the sum of $R_{auto}$ ($R_{root}+R_{auto-ag}$, as described above) and components of $NPP$ consistently fell short of $GPP$, particularly in in young stands (range across forest types and ages: 0.9-7.6 Mg C ha^-1^ yr^-1^).
Moreover, there was not consistent budget closure among the components of $NPP$, and substantially different age trends resulting from the sum of components versus total $NPP$ (Figs. `r TrB_age_trends`-`r BoN_age_trends`).
ALthough age trends of young forests often converged towards mature forest averages, there were also some  discrepancies between young forest trends and mature forest averages (Figs. `r flux_age_trends`, `r TrB_age_trends`-`r BoN_age_trends`, `r all_age_trends`), most notably including a tendency for higher fluxes in regrowth boreal forests than in their mature counterparts (Fig. `r BoN_age_trends`).

@@

*ForC* v3.0 provided unprecedented coverage of most major variables, yielding a broad picture of C cycling in the world's major forest biomes.
Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
Specifically, the major C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs. `r flux_age_trends`- `r stock_age_trends`).
In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 2-5, `r stock_age_trends`).
The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Table 1; Figs. `r flux_age_trends`- `r BoN_age_trends`, `r all_age_trends`).
Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests are correlated with a different set of factors.
Together, these results refine and expand out understanding of C cycling in mature forests, while providing the first global-scale analysis of age trends in multiple forest C cycling stocks and fluxes (Figs. `r TrB_age_trends`-`r BoN_age_trends`).

@@

Our study reveals that most C fluxes quickly increase to a plateau as stands age (Figs. `r flux_age_trends`, `r TrB_age_trends`-`r BoN_age_trends`), consistent with current understanding of age trends in forest C cycling [Fig. `r schematic`; e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_{auto}$) would be minimal immediately following a stand-clearing disturbance.
These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. `r stock_age_trends`).
In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2016; @bond-lamberty_contribution_2004].
In this study, we detect no significant age trends in either variable.

>>>>

#figures
schematic =1
schematic_fluxes_biome="1a"
  schematic_fluxes_age="1b"
  schematic_stocks_biome="1c"
  schematic_stocks_age="1d"
  schematic_fs_age = "1b,d"
figure_map=2
diagram_TrB_mature= 3
diagram_TeB_mature=4
diagram_TeN_mature=5
diagram_BoN_mature=6
all_diagrams_mature="3-6"
flux_age_trends=7
stock_age_trends=8
age_trends_all=9

@@

age_trends_all_caption= "Figure 9 | Age trends in C cycling. Selection of variables for plotting seeks to maximize sample size and broad geographic representation while representing all elements of C cycle. Asterisks indicate variables whose age trends were calculated based on other variables (* young and mature forests; ** young forests only; *** mature forests only), as follows. For all forests: $B_{ag-wood}=max(0,B_{ag}-B_{foliage})$, $B_{root-coarse}=max(0,B_{root}-B_{root-fine})$, $DW_{standing}=max(0, DW_{tot}-DW_{down})$. For tropical forests: $ANPP_{woody}=max(0,ANPP-ANPP_{foliage})$, $R_{auto-ag}=R_{auto}-R{root}$, where $R_{auto}=NPP(1/CUE-1)$ and $CUE$=0.46 (Collati et al. 2020). for non-tropical forests: $ANPP_{woody}=min(ANPP_{stem},ANPP_{woody})$, $R_{auto-ag}=R_{eco}-R{soil}$, . Note that there remain substantial uncertainties as to the functional form of age trends and discrepencies in closure among related variables."

@@

C fluxes commonly increased significantly with stand age (Tables 1, S2, Figs. `r flux_age_trends`, `r age_trends_all`, `r all_age_trends`).
*ForC* contained 16 C flux variables with sufficient data for analyses of age trends in young forests (see Methods).
Of these, ten increased significantly with $log10[age]$: $NEP$, $GPP$, $ANPP$, $ANPP_{woody}$, $ANPP_{foliage}$, $ANPP_{litterfal}$, $BNPP$, $BNPP_{fine}$, $R_{eco}$, and $R_{root}$.
The remaining six--$NPP$, $ANPP_{stem}$, $ANPP_{branch}$, $BNPP_{coarse}$, $R_{soil}$, and $R_{het-soil}$-–displayed no significant relationship to stand age.

```{r eval = generate_figures, echo=FALSE, fig.align = "center", fig.cap =age_trends_all_caption, out.width='100%', fig.pos='H'}

  knitr::include_graphics("tables_figures/stacked_age_trends/age_trends_all.png")
```

@@

"Closure" and internal consistency of the C flux budget was less successful for young than mature forests (Figs. `r age_trends_all`).
Summed regression equations for $R_{soil-het}$ and $R_{root}$ were generally very close to $R_{soil}$.
We calculated $R_{auto-ag}$ as the difference between $R_{eco}$ and $R_{soil}$ (except for tropical forests, which had insufficient $R_{eco}$ data), effectively guaranteeing near-closure of the CO~2~ efflux (respiration) portion of the budget (negative values in Figs. `r age_trends_all`).
In contrast, the  CO~2~ influx portion of the budget generally did not "close": the sum of $R_{auto}$ ($R_{root}+R_{auto-ag}$, as described above) and components of $NPP$ consistently fell short of $GPP$, particularly in in young stands (range across forest types and ages: 0.9-7.6 Mg C ha^-1^ yr^-1^).
Moreover, there was not consistent budget closure among the components of $NPP$, and substantially different age trends resulting from the sum of components versus total $NPP$ (Figs. `r age_trends_all`).
ALthough age trends of young forests often converged towards mature forest averages, there were also some  discrepancies between young forest trends and mature forest averages (Figs. `r flux_age_trends`, `r age_trends_all`, `r all_age_trends`), most notably including a tendency for higher fluxes in regrowth boreal forests than in their mature counterparts (Fig. `r age_trends_all`).

@@

*ForC* v3.0 provided unprecedented coverage of most major variables, yielding a broad picture of C cycling in the world's major forest biomes.
Carbon cycling rates generally increased from boreal to tropical regions and with stand age.
Specifically, the major C fluxes were highest in tropical forests, intermediate in temperate (broadleaf or conifer) forests, and lowest in boreal forests -- a pattern that generally held for regrowth as well as mature forests (Figs. `r flux_age_trends`- `r stock_age_trends`).
In contrast to C fluxes, there was little directional variation in mature forest C stocks across biomes (Figs. 2-5, `r stock_age_trends`).
The majority of flux variables, together with most live biomass pools, increased significantly with stand age (Table 1; Figs. `r flux_age_trends`- `r age_trends_all`, `r all_age_trends`).
Together, these results indicate that, moving from cold to tropical climates and from young to old stands, there is a general acceleration of C cycling, whereas C stocks and $NEP$ of mature forests are correlated with a different set of factors.
Together, these results refine and expand out understanding of C cycling in mature forests, while providing the first global-scale analysis of age trends in multiple forest C cycling stocks and fluxes (Figs. `r age_trends_all`).

@@

Our study reveals that most C fluxes quickly increase to a plateau as stands age (Figs. `r flux_age_trends`, `r age_trends_all`), consistent with current understanding of age trends in forest C cycling [Fig. `r schematic`; e.g., @andersonteixeira_altered_2013; @amiro_ecosystem_2010; @magnani_human_2007].
While limited records in very young (*i.e.*, <5 year old) stands resulted in poor resolution of the earliest phases of this increase for many variables (sometimes detecting no age trend; Table 1), any autotrophic C flux (e.g., $GPP$, $NPP$ and its components, $R_{auto}$) would be minimal immediately following a stand-clearing disturbance.
These would be expected to increase rapidly with the most metabolically active components of biomass, foliage and fine roots, which also increase rapidly with stand age (Fig. `r stock_age_trends`).
In contrast, soil heterotrophic respiration ($R_{het-soil}$) and total soil respiration ($R_{soil}$) are expected to be non-zero following stand-clearing disturbance, although these may decrease with a reduction of root respiration ($R_{soil}$ only) and C exudates or increase in response to an influx of dead roots and litter [@ribeiro-kumara_how_2020; @maurer_carbon_2016; @bond-lamberty_contribution_2004].
In this study, we detect no significant age trends in either variable.
